Study of Intra- and Inter-Job Interference with Different Job Allocations

### 2016-07-11 app_study.tex
some minor tweaks

A parallel application usually conforms to a combination of several basic communication patterns~\cite{roth}. 
At its different execution phases, 
the application's communication behavior may follow certain basic patterns respectively. 
There are many profiling tools available to capture 
information regarding communication patterns of parallel applications~\cite{tau,mpip,sst,oxbow}. 
They can provide information such as the percentage of different MPI operations, 
communication topology, the amount of data transferred between processes, etc.

There are many applications running on HPC systems such as Mira~\cite{bgq} and Titan~\cite{titan}. 
These applications are communication intensive and conform to various communication patterns. 
In this work, we select three representative applications from the DOE Design Forward Project. 
Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications. 
We believe that the communication patterns of these applications 
are representative of a wide array of applications running on leadership-class machines. 
Specifically, we study the Algebraic MultiGrid Solver (AMG), 
Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. 
The communication matrix presented in Figure~\ref{fig:apps_communication_matrix}
are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.

@@

Figure~\ref{fig:amg-communication-topology} shows the communication matrix 
of a small scale AMG execution with 216 MPI Ranks. 
Note that the dominant communication pattern of the application doesn't change with scale. 
We can make the observation that AMG's dominant communication pattern is 3D Nearest Neighbor, 
each rank has intensive communication with up to six neighbors, 
depending on the boundaries of the ranks. 
Applications with similar pattern include PARTISN~\cite{partisn} and SNAP~\cite{snap}.

@@

The second MiniApp covered is CrystalRouter, 
the extracted communication kernel of the full application Nek5000~\cite{nek5000}, 
which is a spectral element CFD application developed at Argonne National Laboratory. 
It features spectral element multi-grid solvers coupled with a highly scalable, 
parallel coarse-grid solver that is widely used for projects including ocean current modeling, 
thermal hydraulics of reactor cores, and spatiotemporal chaos. 
CrystalRouter demonstrates the ``many-to-many'' communication pattern 
through a scalable multi-stage communication process. 
The way CrystalRouter works is the nodes compute for a while, 
then synchronize and communicate, continually alternating between these two types of activities.

@@

Figure~\ref{fig:mg-communication-topology} shows the communication matrix of MultiGrid with 125 ranks. 
We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG. 
However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, 
challenging the maximization of communication locality with respect to ranks. 
In this sense, it can be considered a ``many-to-many'' pattern. 
Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in~\cite{boxlib}.

>>>>

A parallel application usually conforms to a combination of several basic communication patterns~\cite{roth}. 
At its different execution phases, 
the application's communication behavior may follow certain basic pattern respectively. 
There are many profiling tools available to capture 
information regarding communication patterns of parallel applications~\cite{tau,mpip,sst,oxbow}. 
They can provide information such as the percentage of different MPI operations, 
communication topology, the amount of data transferred between processes, etc.

There are many applications running on HPC systems such as Mira~\cite{bgq} and Titan~\cite{titan}. 
These applications are communication intensive and conform to various communication patterns. 
In this work, we select three representative applications from the DOE Design Forward Project. 
Each application exhibits a distinctive communication pattern that is commonly seen in HPC applications. 
We believe that the communication patterns of these applications 
are representative of a wide array of applications running on leadership-class machines. 
Specifically, we study the Algebraic MultiGrid Solver (AMG), 
Geometric MultiGrid (MultiGrid) and CrystalRouter MiniApps. 
The communication matrix of each application is presented in Figure~\ref{fig:apps_communication_matrix}. 
%are generated with the IPM~\cite{ipm} data gathered from publicly available traces~\cite{designforwardwebpage}.

@@

Figure~\ref{fig:amg-communication-topology} shows the communication matrix 
of a small scale AMG execution with 216 MPI ranks. 
Note that the dominant communication pattern of the application doesn't change with scale. 
We can make the observation that AMG's dominant communication pattern is 3D nearest neighbor, 
each rank has intensive communication with up to six neighbors, 
depending on the boundaries of the ranks. 
Applications with similar pattern include PARTISN~\cite{partisn} and SNAP~\cite{snap}.

@@

The second MiniApp covered is CrystalRouter, 
the extracted communication kernel of the full application Nek5000~\cite{nek5000}, 
which is a spectral element CFD application developed at Argonne National Laboratory. 
It features spectral element multi-grid solvers coupled with a highly scalable, 
parallel coarse-grid solver that is widely used for projects including ocean current modeling, 
thermal hydraulics of reactor cores, and spatiotemporal chaos. 
CrystalRouter demonstrates the many-to-many communication pattern 
through a scalable multi-stage communication process. 
The way CrystalRouter works is the nodes compute for a while, 
then synchronize and communicate, continually alternating between these two types of activities.

@@

Figure~\ref{fig:mg-communication-topology} shows the communication matrix of MultiGrid with 125 ranks. 
We can see intensive communication along the diagonal that resembles nearest neighbor communication, similar to AMG. 
However, the communication topology leads to a greater ``spread'' of communication across the set of ranks, 
challenging the maximization of communication locality with respect to ranks. 
In this sense, it can be considered as many-to-many pattern. 
Applications with such similar dominant communication pattern are like FillBoundary, another PDE solver code in~\cite{boxlib}.

### 2016-07-11 conclusion.tex
some minor tweaks

\section{Conclusions}
\label{sec:conclusion}
In this work, we have studied the communication behavior of three parallel applications, namely AMG, CrystalRouter and MultiGrid. Each application has distinctive communication pattern, that can be representative for a group of jobs in HPC workload. We have used a sophisticate simulation tool named CODES from Argonne National Laboratory to simulate the running of these three parallel applications on torus network. The torus network provided by CODES has good fidelity and scalability. We have analyzed the performance of each application's communication in terms of data transfer time by simulating them running on torus networks with different bandwidth and dimensionality configurations. We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.

We have analyzed the intra- and inter-job interference by simulating three applications running on 3D torus network. Based on our comprehensive experiments, we got five observations. 1) The compact allocation may not guarantee the best performance for every application. 2)The applications dominated with Nearest Neighbor communications exhibit relatively stable performance under different allocation shapes, while the applications dominated with Many-to-Many communication exhibit better performance with compact allocation (e.g. 3D balanced). 3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given. 4) An optimal size for allocation units should be determined according to an application's dominant communication pattern. In general, a unit size should be large enough to preserve neighboring communication of the application. 5) Inter-job interference is inevitable in non-contiguous allocation. However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.

We believe that our finding in this work can provide valuable guidance for HPC resource management to make flexible job allocations. Rather than using pre-defined partitions and non-contiguous allocation, future HPC systems should assign each job with preferable resources based on job's communication pattern.

>>>>

\section{Conclusions}
\label{sec:conclusion}
In this work, we have studied the communication behavior of three parallel applications, 
namely AMG, CrystalRouter and MultiGrid. 
Each application has distinctive communication pattern, 
that can be representative for a group of jobs in HPC workload. 
We have used a sophisticated simulation tool named CODES from Argonne National Laboratory 
to simulate the running of these three parallel applications on a torus network. 
The torus network provided by CODES has good fidelity and scalability. 
%We have analyzed the performance of each application's communication in terms of data transfer time 
%by simulating them running on torus networks with different bandwidth and dimensionality configurations. We have found that higher dimensionality of torus network would improve the performance of application with ``Many-to-Many" communication patterns, while application with intensive local communication like ``Nearest Neighbor" won't benefit much from higher dimensionality.

We have analyzed the intra- and inter-job interference by simulating three applications running on the torus network. 
Based on our comprehensive experiments, we got five observations. 
1) The compact allocation may not guarantee the best performance for every application. 
2) The applications dominated with nearest neighbor communications exhibit 
relatively stable performance under different allocation shapes, 
while the applications dominated with many-to-many communication exhibit 
better performance with compact allocation (e.g. 3D balanced). 
3) A good rank-to-node mapping strategy can greatly improve application's performance when specific allocation is given. 
4) An optimal size for allocation units should be determined according to 
an application's dominant communication pattern. 
In general, a unit size should be large enough to preserve neighboring communication of the application. 
5) Inter-job interference is inevitable in non-contiguous allocation. 
However, choosing the proper allocation unit size with job's communication pattern awareness can alleviate such negative effect.

We believe that the findings in this work can provide valuable guidance 
for HPC batch scheduler and resource manager to make flexible job allocations. 
Rather than using pre-defined partitions and non-contiguous placement policy, 
future HPC systems should assign each job with preferable resources based on job's communication pattern.

### 2016-07-11 discussion.tex
some minor tweaks

When an application with dominant communication that is intensive ``Many-to-Many", 
the scheduler should grant it with compact node allocation and exclusive network provision. 
The compact allocation can guarantee the shortest pair-wise distance between all the ranks, 
and make the data transfer between ranks take less hops. 
On the other hand, 
the exclusive network provision will prevent other adjacent applications from sharing network resources, 
thus eliminate the performance degradation due to interference.

Not every application's preferable resource are compact node allocation 
and exclusive network provision. 
As we demonstrate in our analysis, 
applications whose dominant communication pattern contains intensive 
``neighborhood communication" like ``Nearest Neighbor" do not benefit 
from compact node allocation and exclusive network provision. 
Applications, such as AMG, can run with non-contiguous node 
allocation without performance degradation, 
as long as the allocation unit can accommodate their ``neighborhood communication".

The allocation unit size shouldn't be fixed. 
Instead, the scheduler should choose a proper unit size 
based on the scale of ``neighborhood communication" of each job. 
The best unit size should be neither too big nor too small, 
just perfectly fit for the size of that ``neighborhood". 
Big unit size can not be fully utilized and thus cause fragmentation. 
Small unit size will not be able to accommodate the ``neighborhood communication", 
making the intra-job communication less efficient.

>>>>

When an application with dominant communication that is intensive many-to-many, 
the scheduler should provide it with compact node allocation and exclusive network provision. 
The compact allocation can guarantee the shortest pair-wise distance between all the ranks, 
and make the data transfer between ranks take less hops. 
On the other hand, 
the exclusive network provision will prevent other adjacent applications from sharing network resources, 
thus eliminate the performance degradation due to interference.

Not every application's preferable resource are compact allocation 
and exclusive network provision. 
As we demonstrate in our study, 
applications whose dominant communication pattern contains intensive 
``neighborhood communication" like nearest Neighbor do not benefit 
from compact allocation and exclusive network provision. 
Applications, such as AMG, can run with non-contiguous node 
allocation without performance degradation, 
as long as the allocation unit can accommodate their ``neighborhood communication".

The size of allocation unit shouldn't be fixed. 
Instead, the scheduler should choose a proper unit size 
based on the scale of ``neighborhood communication" of each job. 
The best unit size should be neither too big nor too small, 
just perfectly fit for the scale of that ``neighborhood". 
Big unit size can not be fully utilized and thus cause fragmentation. 
Small unit size will not be able to accommodate the ``neighborhood communication", 
making the intra-job communication less efficient.

### 2016-07-11 interference_study.tex
some minor tweaks

In the allocation shapes experiment, 
we select three possible shapes commonly seen on the 3D torus network. 
They are 3D balanced-cube, 3D unbalanced-cube and 2D mesh, 
as shown in Figure \ref{fig:alloc-shapes}.

3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, 
can guarantee the minimum \emph{average pair-wise distance} within the allocation. 
Some research studies~\cite{leung,abhinav-sc13} indicate that 
compact allocation can guarantee jobs with better performance. 
They use a variety of metrics to evaluate the compactness of the allocation, 
such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}. 
In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.

3D unbalanced-cube, shown in green color in Figure \ref{fig:cont_sub1}, 
is a rectangular prism, which is the possible allocation shape 
on the systems with asymmetric networks. 
For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers. 
The network connections in the \emph{y}-direction have 
only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions. 
In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, 
job allocation starts from X-Z plane, 
which leads to a rectangular prism shaped allocation~\cite{RF}.

@@

AMG with 3D ``Nearest Neighbor" communication pattern takes less time 
when being assigned with 3D-unbalanced allocation than 3D-balanced, 
as shown in Figure~\ref{fig:shapstudy-amg}. 
And MultiGrid gets the best performance (shortest data transfer time) 
when running on 3D-balanced allocation, as shown in Figure~\ref{fig:shapstudy-mg}. 
Since MultiGrid's communication pattern is ``Many-to-Many" dominant, 
3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, 
which can reduce the aggregated hops for transferring message among ranks in MultiGrid. 
Since CrystalRouter has both dominant local communication and multi-stage global data transfer, 
the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not 
that obvious as to MultiGrid, as shown in Figure~\ref{fig:shapstudy-cr}.

Lots of research studies have designed complex placement algorithms 
to provide application with the most compact allocation~\cite{leung,LO}. 
However, as shown in our experiments, providing cubic allocation without 
considering application's communication pattern can not guarantee 
the best performance for every application. 
Compact allocation should be provided to applications with intensive global data transfer, 
such as those conforms to ``Many-to-Many" communication pattern. 
Applications with dominant communication pattern like ``Nearest Neighbor" 
can not fully utilize the compact allocation and 
may get even better performance on a relaxed allocation shape.

@@

AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", 
shown in Figure~\ref{fig:diffmap-amg}. 
Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, 
both mapping strategies still preserve the locality of AMG's ``3D Nearest Neighbor" communication pattern. 
The ``Random" mapping totally destroys AMG's communication pattern, 
and results in intra-job interference among all the ranks in AMG. 
The performance degradation caused by using ``Random" mapping is up to 90\%.

The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" 
by reducing the pair-wise distance between ranks, shown in Figure~\ref{fig:diffmap-cr}. 
The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping. 
The ``Random" mapping scatters the ranks within the local neighborhood 
of CrystalRouter and make their communication less efficient.

The ``Cube" mapping benefits MultiGrid's ``Many-to-Many" communication. 
However, due to the small amount of data transfer among ranks in MultiGrid, 
``Cube" mapping fails to exhibit a drastic advantage. 
``Random" mapping neither cause too much degradation for the same reason, 
shown in Figure~\ref{fig:diffmap-mg}.

@@

Inter-job interference has been identified as one of the major culprits 
for application's performance variability~\cite{abhinav-sc13,skinner,rosenthal}. 
Inter-job interference is a more prominent issue for system adopted 
non-contiguous placement policy than systems with contiguous policy. 
Applications communication time have been demonstrated to 
vary from 36\% faster to 69\% slower due to job interference 
when they are running concurrently with non-contiguous allocations~\cite{abhinav-sc13}.

@@

The data transfer time of MultiGrid with different allocation unit sizes doesn't show obvious variability in Figure \ref{fig:interjob-mg}. This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's ``Many-to-Many" pattern. The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, as shown in Figure \ref{fig:interjob-mg} . As the unit size decreases, the data transfer time keep growing, but in a slow way, which means in terms of preserving the locality of MultiGrid, allocation unit of size 16, 8 and 2 serve equally bad.

@@

    \item The applications dominated with ``Nearest Neighbor" communications exhibit 
    relatively stable performance under different allocation shapes, 
    while the applications dominated with ``Many-to-Many" communication exhibit 
    better performance with compact allocation (e.g, 3D balanced).

>>>>

In the allocation shapes experiment, 
we select three possible shapes commonly seen on the 3D torus network. 
They are 3D balanced-cube, 3D unbalanced-cube and 2D mesh, 
as shown in Figure \ref{fig:cont_sub1}.
3D balanced-cube, shown as red in Figure \ref{fig:cont_sub1}, 
can guarantee the minimum \emph{average pair-wise distance} within the allocation. 
Some research studies~\cite{leung,abhinav-sc13} indicate that 
compact allocation can guarantee jobs with better performance. 
They use a variety of metrics to evaluate the compactness of the allocation, 
such as \emph{average pair-wise distance}, \emph{diameter} and \emph{contiguity}. 
In this work, we select 3D balanced-cube as the most compact allocation on a 3D torus network.

3D unbalanced-cube, shown in green color in Figure \ref{fig:cont_sub1}, 
is a rectangular prism, which is a possible allocation shape 
on the systems with asymmetric networks. 
For example, Cray XE6/XK7 systems are deployed 3D tori with Gemini routers. 
The network connections in the \emph{y}-direction have 
only half the bandwidth of the cables used in the \emph{x} and \emph{z} directions. 
In order to take advantage of the faster links in the \emph{x} and \emph{z} directions, 
job allocation starts from X-Z plane, 
which leads to a rectangular prism shaped allocation~\cite{RF}.

@@

AMG with 3D nearest neighbor communication pattern takes less time 
when being assigned with 3D-unbalanced allocation than 3D-balanced, 
as shown in Figure~\ref{fig:shapstudy-amg}. 
And MultiGrid gets the best performance (shortest data transfer time) 
when running on 3D-balanced allocation, as shown in Figure~\ref{fig:shapstudy-mg}. 
Since MultiGrid's communication pattern is many-to-many dominant, 
3D-balanced is the most compact allocation with shortest pair-wise distance between nodes, 
which can reduce the aggregated hops for transferring message among ranks in MultiGrid. 
Since CrystalRouter has both dominant local communication and multi-stage global data transfer, 
the 3D-balance is also the best allocation, but its advantage over 3D-unbalanced is not 
that obvious as to MultiGrid, as shown in Figure~\ref{fig:shapstudy-cr}.

Lots of research studies have designed complex placement algorithms 
to provide applications with the most compact allocation~\cite{leung,LO}. 
However, as shown in our experiments, providing cubic allocation without 
considering application's communication pattern can not guarantee 
the best performance for every application. 
Compact allocation should be provided to applications with intensive global data transfer, 
such as those conforms to many-to-many communication pattern. 
Applications with dominant communication pattern like nearest neighbor 
can not fully utilize the compact allocation and 
may get even better performance on a relaxed allocation shape.

@@

AMG remains roughly the same when it been mapped by ``Linear" and ``Cube", 
shown in Figure~\ref{fig:diffmap-amg}. 
Although ``Linear" and ``Cube" mapping strategies cause some routing overlap for AMG's communication, 
both mapping strategies still preserve the locality of AMG's 3D nearest neighbor communication pattern. 
The ``Random" mapping totally destroys AMG's communication pattern, 
and results in intra-job interference among all the ranks in AMG. 
The performance degradation caused by using ``Random" mapping is up to 90\%.

The ``Cube" mapping improves CrystalRouter's performance over the ``Linear" 
by reducing the pair-wise distance between ranks, shown in Figure~\ref{fig:diffmap-cr}. 
The global data transfer in CrystalRouter will takes fewer hops with ``Cube" mapping. 
The ``Random" mapping scatters the ranks within the local neighborhood 
of CrystalRouter and make their communication less efficient. 
The ``Cube" mapping benefits MultiGrid's many-to-many communication. 
However, due to the small amount of data transfer among ranks in MultiGrid, 
``Cube" mapping fails to exhibit a drastic advantage. 
``Random" mapping neither cause too much degradation for the same reason, 
shown in Figure~\ref{fig:diffmap-mg}.

@@

Inter-job interference has been identified as one of the major culprits 
for application's performance variability~\cite{abhinav-sc13,skinner}. 
Inter-job interference is a more prominent issue for system adopted 
non-contiguous placement policy than systems with contiguous policy. 
Applications communication time have been demonstrated to 
vary from 36\% faster to 69\% slower due to job interference 
when they are running concurrently with non-contiguous allocations~\cite{abhinav-sc13}.

@@

The data transfer time of MultiGrid with different allocation unit sizes 
doesn't show obvious variability in Figure~\ref{fig:interjob-mg}. 
This is because even big allocation unit size like 16 will still fail to preserve MultiGrid's many-to-many pattern. 
The data transfer time is almost doubled when MultiGrid running concurrently with allocation unit size of 16, 
as shown in Figure~\ref{fig:interjob-mg}. 
As the unit size decreases, the data transfer time keep growing, 
but in a slow way, which means in terms of preserving the locality of MultiGrid, 
allocation unit of size 16, 8 and 2 serve equally bad.

@@

    \item The applications dominated with nearest neighbor communications exhibit 
    relatively stable performance under different allocation shapes, 
    while the applications dominated with many-to-many communication exhibit 
    better performance with compact allocation (e.g, 3D balanced).

### 2016-07-11 intro.tex
some minor tweaks

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/goodallocation}
        \caption{Contiguous}
        \label{fig:overview_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/badallocation}
        \caption{Non-contiguous}
        \label{fig:overview_sub2}
    \end{subfigure}%
   \caption{Multiple jobs running concurrently with different allocations. 
   Each job is represented by a specific color. 
   a) shows the effect of contiguous allocation, 
   which could possibly reduce the inter-job interference. 
   b) shows non-contiguous allocation, 
   which may introduce both intra and inter-job interference. 
   }
   \label{fig:overview}
\end{figure}

When submitting jobs to HPC systems, 
users request system resources by specifying the number of compute nodes and expected runtime. 
The batch scheduler is then responsible for dispatching the submitted jobs to the system. 
Typically, there are multiple jobs running concurrently on the system,
resulting in the shared use of resources, particularly network links. 
A prominent problem with this network sharing is the contention over
the network resource among those concurrently running jobs. 
The network contention can cause communication variability and 
performance degradation to those jobs~\cite{abhinav-sc13}. 
This performance degradation can propagate into the queueing time of the following submitted jobs, 
thus leading to low system throughput and utilization~\cite{jose-ipdps15}. 
This adverse effect of network sharing can be mitigated by 
providing jobs with isolated allocations and exclusive network resources.

On the widely used torus-connected HPC systems~\cite{bgq,tofu,titan}, 
two job placement policies are commonly used. 
In \emph{contiguous placement} policy, 
each job gets a compact and contiguous set of computing nodes, 
as shown in Figure~\ref{fig:overview_sub1}. 
The partition-based placement adopted on the Blue Gene 
series systems is an example of contiguous placement policy~\cite{bgloverview}. 
The contiguous placement favors application performance 
through exclusive networking within a compact partition 
and the locality that implies. 
However, contiguous placement can cause both internal fragmentation 
(when more nodes are allocated to a job than it requests) 
and external fragmentation 
(when sufficient nodes are available for a request, but they can not be allocated contiguously), 
therefore leading to lower system utilization than is otherwise possible. 
On the other hand, the \emph{non-contiguous placement} policy, 
adopted by the Cray XT/XE series~\cite{carl-cug}, 
assigns free nodes to jobs regardless of contiguity, 
though of course efforts are made to maximize locality. 
Figure~\ref{fig:overview_sub2} shows the effect of non-contiguous placement. 
While eliminating internal and external fragmentation as seen in contiguous allocation systems, 
in return non-contiguous allocations introduce contention between jobs due to the interleaving of job nodes. 
The non-contiguous job placement can significantly reduce job performance, 
especially for communication-intensive ones~\cite{abhinav-sc13}.

We envision that future HPC systems will be equipped with 
a flexible job placement mechanism which combines the 
best of both contiguous and non-contiguous placement. 
Such a flexible job placement mechanism should take 
shared resource needs (e.g., network resources) of jobs into account 
when making scheduling and allocation decisions. 
With the knowledge and analysis of job communication patterns, 
it can be identified which jobs require exclusive network and allocation locality, and to what degree. 
Then, rather than allocating each job in a ``know-nothing'' manner, 
one may specialize allocation so that, for example, 
only the jobs with stringent network needs are given compact, isolated allocations, 
resulting in maximized utilization and minimized perceivable resource contention effects.

In this work, we focus on an in-depth analysis of intra- and inter-job 
communication interference with different job placement on torus-connected HPC systems. 
Torus-based networks are used on six of the top 10 supercomputers 
on the June 2015 Top500 lists~\cite{top500}. 
The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, 
such as Mira at Argonne Nation Laboratory and 
Sequoia at Lawrence Livermore National Laboratory, 
have their compute nodes connected by a 5D torus network~\cite{bgq}. 
The K computer from Japan uses the ``Tofu'' interconnected system, 
which has a 6D mesh/torus topology~\cite{tofu}. 
Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), 
has nodes connected in a 3D torus within the compute partition~\cite{titan}. 
Although our analyses are based on torus networks, 
the ideas conveyed in this work is applicable to networks with different topologies.

>>>>

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/goodallocation}
        \caption{Contiguous}
        \label{fig:overview_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.22\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/badallocation}
        \caption{Non-contiguous}
        \label{fig:overview_sub2}
    \end{subfigure}%
   \caption{Multiple jobs running concurrently with different allocations. 
   Each job is represented by a specific color. 
   a) shows the effect of contiguous job placement, 
   which could possibly reduce the inter-job interference. 
   b) shows non-contiguous job placement, 
   which may introduce both intra and inter-job interference. 
   }
   \label{fig:overview}
\end{figure}

When submitting jobs to HPC systems, 
users request system resources by specifying the number of compute nodes and expected runtime. 
The batch scheduler is then responsible for dispatching the submitted jobs to the system. 
Typically, there are multiple jobs running concurrently on the system,
resulting in the shared use of resources, particularly network links. 
A prominent problem with this network sharing is the contention over
the network among those concurrently running jobs. 
The network contention can cause communication variability and 
performance degradation to those jobs~\cite{abhinav-sc13}. 
This performance degradation can propagate into the queueing time of the following submitted jobs, 
thus leading to low system throughput and utilization~\cite{jose-ipdps15}. 
This adverse effect of network sharing can be mitigated by 
providing jobs with isolated allocations and exclusive network resources.

On the widely used torus-connected HPC systems~\cite{bgq,tofu,titan}, 
two job placement policies are commonly used. 
In \emph{contiguous placement} policy, 
each job gets a compact and contiguous set of computing nodes, 
as shown in Figure~\ref{fig:overview_sub1}. 
The partition-based placement adopted on the Blue Gene 
series systems is an example of contiguous placement policy~\cite{bgloverview}. 
The contiguous placement favors application performance 
through exclusive networking within a compact partition 
and the locality that implies. 
However, contiguous placement can cause both internal fragmentation 
(when more nodes are allocated to a job than it requests) 
and external fragmentation 
(when sufficient nodes are available for a request, but they can not be allocated contiguously), 
therefore leading to lower system utilization than is otherwise possible. 
On the other hand, the \emph{non-contiguous placement} policy, 
adopted by the Cray XT/XE series~\cite{carl-cug}, 
assigns free nodes to jobs regardless of contiguity, 
though of course efforts are made to maximize locality. 
Figure~\ref{fig:overview_sub2} shows the effect of non-contiguous placement. 
While eliminating internal and external fragmentation as seen in contiguous placement systems, 
in return non-contiguous placement policy introduces contention between jobs due to the interleaving of job nodes. 
The non-contiguous job placement can significantly reduce job performance, 
especially for communication-intensive ones~\cite{abhinav-sc13}.

We envision that future HPC systems will be equipped with 
a flexible job placement mechanism which combines the 
best of both contiguous and non-contiguous policy. 
Such a flexible job placement mechanism should take 
shared resource needs (e.g., network resources) of jobs into account 
when making scheduling and allocation decisions. 
With the knowledge and analysis of job communication patterns, 
it can be identified which jobs require exclusive network and compact allocation, and to what degree. 
Then, rather than allocating each job in a ``know-nothing'' manner, 
one may specialize job placement policy so that, for example, 
only the jobs with stringent network needs are given compact, isolated allocations, 
resulting in maximized utilization and minimized perceivable resource contention effects.

In this work, we focus on an in-depth analysis of intra- and inter-job 
communication interference with different job placements on torus-connected HPC systems. 
Torus-based networks are used on six of the top 10 supercomputers 
on the June 2015 Top500 lists~\cite{top500}. 
The current generation of IBM Blue Gene/Q (BG/Q) supercomputers, 
such as Mira at Argonne Nation Laboratory and 
Sequoia at Lawrence Livermore National Laboratory, 
have their compute nodes connected by a 5D torus network~\cite{bgq}. 
The K computer from Japan uses the ``Tofu'' interconnected system, 
which has a 6D mesh/torus topology~\cite{tofu}. 
Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), 
has nodes connected in a 3D torus within the compute partition~\cite{titan}. 
Although our analyses are based on torus networks, 
the ideas conveyed in this work is applicable to networks with different topologies.

### 2016-07-11 related_work.tex
some minor tweaks

%\textbf{P3: Interference between concurrently running jobs}\\
The interference among concurrently running jobs on HPC systems 
have been identified as major culprit for job's performance variability. 
Bhatel et al. found that concurrently running applications interfere with each other, 
and cause their communication time varied from 
36\% faster to 69\% slower on different HPC systems~\cite{abhinav-sc13}. 
Skinner et al. found that  there is a 2-3$\times$ slowdown in MPI$\textunderscore$Allreduce 
due to network contention from other concurrently running jobs~\cite{skinner}. 
Rosenthal et al. found increasing network bandwidth provide limited benefits to a number of applications. 
The applications that send mostly small messages or large messages asynchronously 
are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth~\cite{rosenthal}.

%\textbf{P4: Job allocation(little bit)}\\
Several research studies focus on optimizing job allocation on HPC systems 
to alleviate the interference between concurrently running jobs. 
Hoefler et al. proposed to use performance modeling techniques to analyze factors that 
impact the performance of parallel scientific applications~\cite{hoefler-modeling}. 
However, as the scale of HPC systems continue to grow, 
the interference of concurrently running jobs is getting worse, 
which is hard to be quantified by performance profiling tools alone. 
Bogdan et al. provide a set of guidelines of how to configure a network with Dragonfly topology 
for workload with Nearest Neighbor communication pattern~\cite{Bogdan-hpdc14}. 
Dong et al. have developed simple benchmarks that conforms to four different communication patterns, 
namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, 
to demonstrate the effectiveness of this highly parallel 5D torus network~\cite{Dong-SC11}.

Our work is different from all these works in the following ways. 
First, we focus on the dominant communication patterns rather than any specific application. 
We believe this can provide a guideline for other research work. 
Secondly, we explore the intra- and inter-job interference between concurrently running jobs, 
while similar work such as~\cite{abhinav-sc13} only focuses 
on the single application's performance degradation due to network contention. 
Finally, we analyze the impact of different allocation strategies to job's communication behaviors. 
We further identify the optimal allocation strategy for 
each application with a specific dominant communication pattern. 
Based on our study, we claim that further batch scheduler should take 
job's communication pattern into consideration for its placement decision making.

>>>>

%\textbf{P3: Interference between concurrently running jobs}\\
The interference among concurrently running jobs on HPC systems 
have been identified as major culprit for job's performance variability. 
Bhatele et al. found that concurrently running applications interfere with each other, 
and cause their communication time varied from 
36\% faster to 69\% slower on different HPC systems~\cite{abhinav-sc13}. 
Skinner et al. found that  there is a 2-3 times of slowdown in MPI$\textunderscore$Allreduce 
due to network contention from other concurrently running jobs~\cite{skinner}. 
%Rosenthal et al. found increasing network bandwidth provide limited benefits to a number of applications. 
%The applications that send mostly small messages or large messages asynchronously 
%are not bandwidth bounded, hence, benefit only slightly or not at all from increased bandwidth~\cite{rosenthal}.

%\textbf{P4: Job allocation(little bit)}\\
Several research studies focus on optimizing job allocation on HPC systems 
to alleviate the interference between concurrently running jobs. 
Hoefler et al. proposed to use performance modeling techniques to analyze factors that 
impact the performance of parallel scientific applications~\cite{hoefler-modeling}. 
However, as the scale of HPC systems continue to grow, 
the interference of concurrently running jobs is getting worse, 
which is hard to be quantified by performance profiling tools alone. 
Bogdan et al. provide a set of guidelines of how to configure a Dragonfly network 
for workload with nearest neighbor communication pattern~\cite{Bogdan-hpdc14}. 
Dong et al. have developed simple benchmarks that conforms to four different communication patterns, 
namely ping-pong, nearest neighbor, broadcast and all$\textunderscore$reduce, 
to demonstrate the effectiveness of this highly parallel 5D torus network~\cite{Dong-SC11}.

Our work is different from all these works in the following ways. 
First, we focus on the dominant communication patterns rather than any specific application. 
We believe this can provide a guideline for other research work. 
Secondly, we explore the intra- and inter-job interference between concurrently running jobs, 
while similar work such as~\cite{abhinav-sc13} only focuses 
on the single application's performance degradation due to network contention. 
Finally, we analyze the impact of different placement strategies to job's communication behaviors. 
We further identify the optimal placement strategy for 
each application with a specific dominant communication pattern. 
Based on our study, we claim that further batch scheduler should take 
job's communication pattern into consideration for its placement decision making.
